---
title: "Meta-Analysis Workflow"
author: "Charles T. Gray, Malcolm Barrett, W. Kyle Hamilton"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r warning=FALSE}
library(metaverse)
library(mbmisc)
```


Model objects can be challenging to work with in their raw form in R. The [`broom::`](https://broom.tidyverse.org/) package provides tools to extract commonly desired model output in **table** form:

> 1. tidy() summarizes information about model components
2. glance() reports information about the entire model
3. augment() adds information about observations to a dataset ()

## demonstration

We begin by creating a random-effects model object via `metafor::`, which is included in the `metaverse::`. 

```{r data}
# grab a dataset from metafor
# from the metafor homepage
dat.bcg <- metafor::dat.bcg
ma_model <-
  escalc(
    measure = "RR",
    ai = tpos,
    bi = tneg,
    ci = cpos,
    di = cneg,
    data = dat.bcg
  ) %>% rma(yi, vi, data = ., method = "EB")

```

### broom functionality

We can now use `broom::` functionality to manage the output in a common format

> 1. tidy() summarizes information about model components

```{r tidy}
ma_model %>% tidy() %>% knitr::kable()

```

> 2. glance() reports information about the entire model

```{r glance}
ma_model %>% glance()  %>% knitr::kable()
```

> 3. augment() adds information about observations to a dataset ()

```{r}
ma_model %>% augment() %>% knitr::kable()
```

# Analysis

We can use functions from `tidymeta::` to conduct subgroup meta-analyses

```{r}
ma_model_wkh <- metafor::dat.bcg


ma <- ma_model_wkh %>% 
  escalc(
    measure = "RR",
    ai = tpos,
    bi = tneg,
    ci = cpos,
    di = cneg,
    data = .) %>%
  dplyr::group_by(alloc) %>% 
  meta_analysis(yi = yi, vi = vi, slab = author, exponentiate = TRUE) 

ma

```
## Forest Plot

`tidymeta::` also has some handy ggplot based graphics for meta-analysis

```{r warning=FALSE}
fp <- ma %>% 
  forest_plot(group = alloc)

fp
```

## Publication Bias
Estimates the Vevea and Hedges (1995) weight-function model using the `pubBias::` package.

```{r warning=FALSE}
pubBias::weightrMeta(ma)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# 
# 
#    MM                   
#   <' \___/|            
#     \_  _/  
#       ][  
#    Chicken 
#                          
```

