---
title: "extracting common information from meta-analysis model output"
author: "Charles T. Gray and Malcolm Barrett"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r }
library(metaverse)
```


Model objects can be challenging to work with in their raw form in R. The [`broom::`](https://broom.tidyverse.org/) package provides tools to extract commonly desired model output in **table** form:

> 1. tidy() summarizes information about model components
2. glance() reports information about the entire model
3. augment() adds information about observations to a dataset ()

## demonstration

We begin by creating a random-effects model object via `metafor::`, which is included in the `metaverse::`. 

```{r data}
# grab a dataset from metafor
# from the metafor homepage
ma_model <-
  escalc(
    measure = "RR",
    ai = tpos,
    bi = tneg,
    ci = cpos,
    di = cneg,
    data = dat.bcg
  ) %>% rma(yi, vi, data = ., method = "EB")

```

### broom functionality

> 1. tidy() summarizes information about model components

```{r tidy}
ma_model %>% tidy()

```

> 2. glance() reports information about the entire model

```{r glance}
ma_model %>% glance()
```

> 3. augment() adds information about observations to a dataset ()

```{r}
ma_model %>% augment()
```


## requesting more features

Currently, this is restricted to `metafor::rma` model objects, but Malcolm and Charles are most happy to collaborate on expanding the functionality to other meta-analytic objects. 

The easiest way to let us know is to log an [issue](https://github.com/rmetaverse/metaverse/issues) on the `metaverse::` GitHub repository.   
